---
title: "Goal programming-Assignment"
author: "Harish Kunaparaju"
date: "2022-11-05"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The Research and Development Division of the Emax Corporation has developed three new 
products. A decision now needs to be made on which mix of these products should be produced. 
Management wants primary consideration given to three factors: total profit, stability in the 
workforce, and achieving an increase in the company’s earnings next year from the $75 million 
achieved this year

#importing libraries
```{r}
library(lpSolve)
library(lpSolveAPI)
library(goalprog)
```

objective function
Maximize Z = P - 6C - 3D, where  
P = total (discounted) profit over the life of the new products, 
C = change (in either direction) in the current level of employment, 
D = decrease (if any) in next year’s earnings from the current year’s level. 


#lptable of each new product shown in the table
```{r}
lp<- matrix(c("Total Profit", "Employment Level", "Earnings Next Year",
                       20,6,8,
                       15,4,7,
                       25,5,5,
                       "Maximize","=50",">=75",
                       "Millions of Dollars", "Hundreds of Employees", "Millions of Dollars"), ncol=6, byrow = F)
colnames(lp) <- c("Factor","Product 1", "Product 2", "Product 3", "Goal", "Units")
as.table(lp)

```


```{r}

lp1<-read.lp("goal.lp")
lp1
```


#goal programming model 
```{r}
solve(lp1)
```
*the return value of 0 indicates that the model was successfully solved.*

#-objective
```{r}
get.objective(lp1)
```
#-variable value
```{r}
get.variables(lp1)
```

#interpretation

1.The units of combination that maximize the objective function are X1, X2, and X3. X1 = Product1, X2 = Product2, and X3 = Product3. But X3 has undergone a shift. The company can only make 15 units of Product 3, which is the only product, in order to maximize profit. The intention was to stabilize employment levels with a cap of 50 hundred employees, however in this case, the firm's employment levels were surpassed by 250 employees, necessitating the payment of a penalty for the excess/rise in the employee count.
 
 # 2.the goal of y2p and y2m was to see the deviation either in positive or negative in the next years earnings from the current level.
 
 # 3.The profit that the firm maximizing is 225 Million Dollars.